Perihelion advance of Mercury

### 2019-06-20 perihelion_advance_of_mercury.Rmd
docs
consistency check with GR=FALSE orbit

In the above diagram we see over 15.4 orbits of Mercury, exhibiting no
visible deviation from an exactly closed orbital path.  Nevertheless,
the orbit is not closed as Mercury has a tiny amount of orbital
precession.  To calculate this precession we need to simulate the
motion for a longer time:

@@

We need to isolate the change points (actually calculating aphelion
rather than perihelion for convenience):

@@

plot(tee,phi-pi,main='orbital angle at aphelion',xlab='time (seconds)',ylab='angle of aphelion')
```

@@

>>>>

In the above diagram we see over 15.4 orbits of Mercury, exhibiting no
visible deviation from an exactly closed orbital path.  Nevertheless,
the orbit is not closed as Mercury has a tiny amount of orbital
precession.  Note that the Schwarzschild radius would be microsopic at
this scale (the size of the central disk is dominated by the width of
the plotting line used for the event horizon).  To calculate this
precession we need to simulate the motion for a longer time:

@@

We need to isolate the change points, here found isolating iterations
$i$ such that ```r[i-1]<r[i]``` and r[i]>r[i+1]```:

@@

plot(tee,phi-pi,main='orbital angle of Mercury at perihelion',xlab='time (seconds)',ylab='angle of perihelion')
```

@@

By way of a consistency check, we can plot a similar diagram but with
the classical approximation:

```{r}
out <- trajectory(merc_perihelion, m=d, h=h, tau=seq(from=0,to=35e16,len=1e7),GR=FALSE)
jj1 <- c(NA,out$rdot)
jj2 <- c(out$rdot,NA)
change <- (jj1>0) & (jj2<0)
tee    <- out$time[change]
phi <- out$phi[change]
phi <- (phi) %% (2*pi)
plot(tee,phi-pi,main='orbital angle at of Mercury at perihelion, classical case',xlab='time (seconds)',ylab='angle of perihelion')
```

In the above plot, note how there is no trend, consistent with the
Newtonian case of zero perihelion advance.
